A Time Varying Vector Error Correction Model Applied to Foreign Exchange Market

### 2016-03-22 articles-conferences/elsevier-journal-2015/src/background.tex
Agregar estos repos a la tesis

\chapter{Machine Learning Models}
Machine learning is a scientific discipline focused on the development of algorithms focused on learning from examples. This idea has become central to the design of search engines, robots systems and forecasts applications which process large data sets. However, machines require an extended training period when developing algorithms to predict future behaviour such as financial time series forecasting where new data is available in a short period of time.  Online machine learning techniques tackle this problem and allow to update the model with new data or to compute a new model using less data allowing to give a response in a short period of time.

\vspace{0.5cm}

\section{Introduction}
Machine Learning (ML) studies computer algorithms for learning
something such as to complete a task, to make accurate predictions or to behave intelligently. The learning is always based on samples and the objective is about to do better in the future based on the past experiences in an automatic way. ML is a subarea of artificial intelligence and broadly intersects with other fields such as statistics, mathematics, physics and computer science. 
There are many examples of machine learning problems: time series forecasting, image processing, face detection, spam filtering, weather prediction, search engines, among many others.

ML models are often more accurate than what can be created through direct programming. The reason is that ML models are data driven and are able to examine large amounts of data.

There are three types of ML classified depending on the nature of the learning input or output available to a learning system.
\begin{description}
\item[supervised learning]  the input data is a tuple which contains the example input and its desired outputs (also called labels). The list of tuples is called the training set. The concept of supervised learning comes from the supervisor, acting as a teacher in the learning process. The goal is to learn a general rule that maps inputs to outputs optimising a target function. There are two related problem types in supervised learning: classification and regression problems \cite{bishop2006}. Its two mainstream approaches are: support vector machines (SVMs) \cite{vapnik1998} and ensemble learning \cite{breiman1998}. Furthermore supervised learning can be categorised into offline or batch learning and online learning (see section \ref{sec:onoffline}). 
\item[unsupervised learning] also known as clustering \cite{ben2005}. In this type of learning no labels are given and the system has to find a structure on its own, discovering hidden patterns in data.
\item[reinforcement learning] is the problem faced by an agent that must learn 
through trial-and-error interactions with a dynamic environment. It is based on programming agents by reward and punishment without needing to specify how the task is to be achieved \cite{sutton1998}.
\end{description}

%%% Learning problem. Ac;a ma;ana, formulas y descripcion de tallada del problema de aprendizaje

\section{Statistical learning theory} \label{sec:mltheory}

All ML problems can be viewed as optimisation problems. The ML core task is to define a learning criterion, i.e the function to be optimised.

In this thesis we will emphasised the supervised learning problem since it is the most common way of modelling financial problems.

Supervised learning consists in to find a learning function $f: X \rightarrow Y$ which models a set of examples called training set $S$ which have been drawn randomly and independently according to a unknown joint distribution function $p(x,y)$ on $X \times Y$:

$$
S= \{(x_1,y_1),\,(x_2,y_2),\dots\,,\,(x_n,y_n)\} \in \mathcal{S}, (x_i,y_i) \in X\times Y \quad \forall i = 1\dots n
$$

\noindent where $X \subseteq \mathbb{R}^m$ and $Y = \mathbb{R}$ in case of regression problems and $Y = \{1,-1\}$  if it is a classification problem.

A learning algorithm $\mathcal{A}$ takes as input a data set $S \in \mathcal{S}$ and output a function $f_S$:

\begin{eqnarray*}
\mathcal{A}: \mathcal{S} &\rightarrow & \mathcal{H} \\
S &\rightarrow &\mathcal{A}(S) = f_S
\end{eqnarray*}

\noindent where $\mathcal{H}$ is called the hypothesis space, is the space of functions that the algorithm is allowed to search. The selection of $f_S$ is based on a loss function $V(f(x),y)$ which expected risk has to be minimised.

\[
E[V(f(x),y)] = \int V(f(x),y) dp(x,y)
\]

\noindent $V(f(x),y)$ denote the price paid for mistakes. Therefore, $V(f(x),y)=0$ if$f(x)=y$.

For regression, the most common loss function is square loss or L2 loss function:
\begin{equation}
\label{eq:l2loss}
V(f(x),y) = (f(x)-y)^2
\end{equation}
\noindent another option is the L1 loss
\begin{equation}
\label{eq:l1loss}
V(f(x),y) = |f(x)-y| \, .
\end{equation}
\noindent the choice of loss function here gives rise to several well-known learning algorithms such as regularised least squares and support vector machines. %verificar esto

Since the true distribution is unknown and only training samples are available, the objective is to estimate a function $\hat{f}$ through empirical risk (training error) minimisation (ERM):

\begin{equation} 
\label{eq:erm}
R_{\text{emp}}[f] = \frac{1}{n} \sum_{i=1}^n V(f(x),y)
\end{equation}

\subsection{Learning algorithms}
In all learning algorithms that are trained from example data, there is a tradeoff \cite{dietterich2003} between three factors:
\begin{description}
\item[Complexity of the hypothesis class] for best generalisation, we should adjust the complexity of the learner model $\hat{f}$ to the complexity of the data $f$. In polynomial regression, the complexity parameter is the order of the fitted polynomial, and therefore we need to find a way to choose the best order that minimises the generalisation error, that is, tune the complexity of the model to best fit the complexity of the function inherent in the data.
\item[Generalisation accuracy on new examples] is the capability of the algorithm to generate the right output for an input instance outside the training set.
\item[The amount of training data]  generally generalisation accuracy increases as the amount of training data increases.
\end{description}

Very complex models will fit the training data better (low bias) but it could overfit it and generalise poorly (high variance), this is also called the bias-variance tradeoff.  Generalisation accuracy can sometimes be improved with more training data. The figure \ref{fig:tradeoff} illustrates how generalisation accuracy depends on the complexity of the model and the amount of training data.

\begin{figure}[!h]
  \centering
  \includegraphics[width=0.8\textwidth]{img/3tradeoff}
  \caption{Triple tradeoff in empirical learning}
  \label{fig:tradeoff}
\end{figure}

\subsection{The bias-variance tradeoff} \label{sec:biasvar}

Models learning error can be split into two main components: error due to bias and error due to variance. Bias measures how far off in general these models' predictions are from the correct value. Variance is taken as the variability of a model prediction for a given data point.

If we have a learning model $\hat{f}$, its expected generalisation error on an unseen or testing sample $x$ can be decomposed as follows:

\begin{align}
\label{eq:bvtrade}
\mathrm{E}\Big[\big(y - \hat{f}(x)\big)^2\Big]
 & = \mathrm{Bias}(\hat{f}(x))^2 + \mathrm{Var}(\hat{f}(x)) + \sigma^2
\end{align}
\noindent where:
\begin{align*}
 \mathrm{Bias}(\hat{f}(x)) &= \mathrm{E}\big[\hat{f}(x)\big] - f(x) \\
 \mathrm{Var}(\hat{f}(x)) &= \mathrm{E}\Big[ \big( \hat{f}(x) - \mathrm{E}[\hat{f}(x)] \big)^2 \Big] 
\end{align*}

@@

\begin{eqnarray*}
E[(y-\hat{f}(\mathbf{x}))^2] & =&
E[(y-f(\mathbf{x})+f(\mathbf{x})-\hat{f}(\mathbf{x}))^2] \\
&=& E[(B-f(\mathbf{x}))^2] +
E[(f(\mathbf{x})-\hat{f}(\mathbf{x}))^2] + \dots \\
& &
2\cancelto{0}{E[y-f(\mathbf{x})]}E[f(\mathbf{x})-\hat{f}(\mathbf{x})] \\
&=& \sigma^2 + MSE(\hat{f}(\mathbf{x}))
\end{eqnarray*}

\begin{eqnarray*}
    MSE(\hat{f}(\mathbf{x})) &=&
    E[f(\mathbf{x})-\hat{f}(\mathbf{x}))]^2 \\
    &=& E[f(\mathbf{x})-E[\hat{f}(\mathbf{x})] +
    E[\hat{f}(\mathbf{x})]-\hat{f}(\mathbf{x}))]^2 \\
    &=&  E[(f(\mathbf{x})-E[\hat{f}(\mathbf{x})])^2] +
      E[(\hat{f}(\mathbf{x})-E[\hat{f}(\mathbf{x})])^2] + \dots\\
    & & E[(f(\mathbf{x})-E[\hat{f}(\mathbf{x})])] 
    \cancelto{0}{E[(\hat{f}(\mathbf{x})-E[\hat{f}(\mathbf{x})])]} \\
    &=& Bias(\hat{f}(\mathbf{x}))^2 + Var(\hat{f}(\mathbf{x}))
\end{eqnarray*}

$\blacksquare$

Bias is introduced by the model selection. Therefore the model building process is repeated (through resampling) and substantially different averages of prediction values are obtained, bias will be high. The error due to variance is the amount by which the prediction, over one training set, differs from the expected predicted value, over all the training sets. Variance measures how inconsistent are the predictions from one another, over different training sets, not whether they are accurate or not.

This generalisation error is unknown, but it can be estimated using a training sample, a testing sample or by estimating the model complexity. Figure \ref{fig:traintesterror} shows prediction error for training and testing sample.  In general, in the training sample prediction error is decreasing with more complex models but it overfit the data and it is not a good model for the test data. Figure \ref{fig:bvtradeoff} shows how testing error can be decomposed into bias and variance components. The best model will be the one has a balance between bias and variance.

\begin{figure}[!h]
  \centering
  \includegraphics[width=0.8\textwidth]{img/model_complexity}
  \caption{Training and test error}
  \label{fig:traintesterror}
\end{figure}

The objective is to simultaneously reduce bias and variance as much as possible in order to obtain as accurate model as is feasible.  However, there is a tradeoff to be made when selecting models. Models that exhibit small variance and high bias underfit the truth target.  Models that exhibit high variance and low bias overfit the truth target.

This bias-variance tradeoff is the reason why in order to obtain the best model, data is usually broken in three subsets: training set, validation set and testing set. Training set is used to determine the model, validation set is used to estimate the generalisation error and finally testing set is used to estimate the accuracy of the model. Usually the partition is 50\% for training set, 25\% for validation and 25\% for testing purposes. This procedure can be extended to a Cross Validation (CV) procedure \cite{geisser1975}, usually used when the amount of data is limited. Various splitting strategies lead to various CV estimates. In K-fold cross-validation the training data is divided randomly into K distinct subsets, then the network is trained using K-1 subsets, and tested on the remaining subset. The process of training and testing is then repeated for each of the K possible choices of the subset omitted from the training. The average performance on the K omitted subsets is then our estimate of the generalisation performance.

\begin{figure}[!h]
  \centering
  \includegraphics[width=0.8\textwidth]{img/biasvariancetradeoff}
  \caption{Bias variance tradeoff}
  \label{fig:bvtradeoff}
\end{figure}

Another way to control the complexity of the model is to include in the optimisation function not only the generalisation error but a penalisation of high complex models.

\subsection{Regularization}
Since ERM is an ill-posed problem, Tikhonov introduced a regularisation which ensures well-posedness and generalisation of ERM, i.e prevents overfit, by constraining the hypothesis space $\mathcal{H}$ usually called regularised ERM or Tikhonov regularisation.

Tikhonov regularisation minimised over the hypothesis space $\mathcal{H}$ for a fixed positive parameter $\lambda$ the following:

\begin{equation} 
\label{eq:rerm}
R_{\text{emp}}[f] = \frac{1}{n} \sum_{i=1}^n V(f(x),y) + \lambda \mathcal{R}(f)
\end{equation}

\noindent where $\mathcal{R}(f)$ is the regulariser, a penalisation on $f$.

\subsection{Ridge Regression}

One example of regularisation in linear models is Ridge Regression (RR), which is a regularised least squares method.
The least squares (LS) method is a well known way to solve a regression
problem. 
LS method consists of minimising the sum of squared errors:

\begin{eqnarray*}
\label{eq:lsm}
 J(\mathbf{X}) =& \displaystyle \sum_{t=1}^N (f(\mathbf{x}_t)-y_t)^2 \\
 =&\displaystyle \sum_{t=1}^N (\mathbf{\mathbf{X}}^\top {\mathbf{x}}_t-y_t)^2 \\
 =& \| \mathbf{A}\mathbf{\mathbf{X}} - \mathbf{Y} \|_2^2 
\end{eqnarray*}

\noindent which is equivalent to find a solution to 
\begin{equation}
\label{eq:regproblem}
\underset{m \times n}{\mathbf{A}} \enskip \underset{n \times
l}{\mathbf{\mathbf{X}}} = \underset{m \times l}{\mathbf{Y}}
\end{equation}

The optimal solution for $\mathbf{\mathbf{X}}$ is:
\begin{equation}
\label{eq:MP}
\mathbf{\mathbf{X}}= \mathbf{A}^+ \mathbf{Y}
\end{equation}

\noindent where $\mathbf{A}^+$ is the Moore-Penrose pseudo-inverse
which can be written as follows:

\begin{equation}
\label{eq:pseudoinverse}
\mathbf{A}^+= (\mathbf{A}^\top \mathbf{A})^{-1}\mathbf{A}^\top \, .
\end{equation}

To solve equation~(\ref{eq:pseudoinverse}) is equivalent to solve the following
optimisation problem:

@@

\begin{equation*}
	\mathbf{C}=
	\begin{bmatrix} \quad &\mathbf{A}& \quad \\ \hdashline \quad
& \sqrt\lambda\mathbb{I}&\quad  
\end{bmatrix} 
\quad \text{and} \quad
	\mathbf{F}=\begin{bmatrix} \quad &\mathbf{Y}& \quad \\ \hdashline \quad
& \mathbf{0}&\quad  \end{bmatrix} \, .
\end{equation*}

Applying equation~(\ref{eq:MP}) and considering that $\mathbf{C}^\top
\mathbf{C} = \mathbf{A}^\top \mathbf{A} + \lambda \mathbb{I}$ and 
$\mathbf{C}^\top \mathbf{F}=\mathbf{A}^\top \mathbf{Y} $ we have:

\begin{eqnarray*}
\mathbf{X}(\lambda)&=&(\mathbf{C}^\top
\mathbf{C})^{-1}\mathbf{C}^\top \mathbf{F} \\
&=& (\mathbf{A}^\top \mathbf{A} + \lambda \mathbb{I})^{-1} \mathbf{A}^\top \mathbf{Y}
\end{eqnarray*}

$\blacksquare$

\begin{equation}
    \label{eq:compactsvd}
    \underset{m \times n}{\mathbf{A}}=
    \underset{m \times k}{\mathbf{U_1}} \enskip
    \underset{k \times k}{\mathbf{\Sigma_1}} \enskip
    \underset{k \times n}{\mathbf{V_1}^\top}
\end{equation}

>>>>

\section{\uppercase{Background}}
\label{sec:background}
\noindent

\subsection{Spurious Regression}

Regression analysis is widely used to test economic theories. But sometimes
regression is done between two variables, lets say $x_t$ and $y_t$, which are
not causally related. This occurs at the presence of a third stochastic variable
$z_t$ which is the true causal of the first two variables $x_t$ and $y_t$. The
regression measure R-squared gives high values in presence of spurious
relationships. This measure would means that the model explains all the
variability of the response data around its mean which is not true.

The R-squared measure is calculated as follows:

\begin{equation}
\label{eq:rsquared}
R^2 = 1 - \frac{\sum_{t=1}^T \epsilon_t^2}{\sum_{t=1}^T(y_t - \bar{y})^2}
\end{equation}

In presence of spurious relationships the denominator of equation
\ref{eq:rsquared} becomes very large because a large weight is placed on extreme
observations on either side of the mean $\bar{y}$.
In order to detect this, other statistic must be considered such as the
Durbin-Watson statistic.

% Acá agregar gráfico de spurious regression

\subsection{Integration and Cointegration}
A time series $\mathbf{y}$ is said to be integrated of order $d$ if after
differentiating the variable $d$ times, we get an I(0) process, more precisely:

\[
(1-L)^d \mathbf{y} \sim \text{I(0)} \, ,
\]

\noindent where I(0) is a stationary time series and $L$ is the lag operator:

\[
(1-L)\mathbf{y} = \Delta \mathbf{y}=\mathbf{y}_t  -\mathbf{y}_{t-1} \quad \forall t
\]

Let $\mathbf{y}_t = \{\mathbf{y}^1, \dots, \mathbf{y}^l\}$ be a set of $l$
stationary time series I(1) which are said to be cointegrated if a vector
$\beta=[\beta(1),\dots,\beta(l)]^\top \in \mathbb{R}^l$  exists such that the
time series,

\begin{equation}
 \mathbf{Z}_t:= \beta^\top \mathbf{y}_t = \beta(1) \mathbf{y}^1 + \dots + \beta(l) \mathbf{y}^l \sim
 \text{I(0)}\, .
\end{equation}

In other words, a set of I(1) variables is said to be cointegrated if
a linear combination of them exists which is I(0).

\subsection{Vector Autorregresive Models}\label{sec:varvec}

VECM is a special case of VAR model and both describe the joint behaviour of a
set of variables.

VAR($p$) model is a general framework to describe the behaviour of a set of $l$
endogenous variables as a linear combination of their last $p$ values. These
$l$ variables at time $t$ are represented by the vector $\mathbf{y}_t$ as
follows:

\begin{equation}
\label{eq:variables}
\mathbf{y}_t = 
\begin{bmatrix} y_{1,t} &
y_{2,t} &
\dots &
y_{l,t}
\end{bmatrix}^\top \, ,
\end{equation}
\noindent where $y_{j,t}$ corresponds to the time series $j$ evaluated at
time $t$.

The VAR(p) model describes the behaviour of a dependent variable in terms
of its own lagged values and the lags of the others variables in the
system. The model with $p$ lags is formulated as the following:

\begin{equation}
\label{eq:var}
 \mathbf{y}_t = \phi_1 \mathbf{y}_{t-1}  + \dots +   \phi_p\mathbf{y}_{t-p}
+ \mathbf{c} + \mathbf{\epsilon}_t \, ,
\end{equation}

\noindent where $t=p+1, \dots N$,  ${\phi_1,\dots,\phi_p}$ are $l \times l$
matrices of real coefficients,
$\mathbf{\epsilon}_{p+1},\dots,\mathbf{\epsilon}_N$ are error terms,
$\mathbf{c}$ is a constant vector and $N$ is the total number of samples.

The VAR matrix form of equation~(\ref{eq:var}) is:

%\begin{equation}\label{eq:vareq}
%\mathbf{B} = \mathbf{A} \mathbf{X} + 
%\mathbf{E} \, , 
%\end{equation}
%
%\noindent where:
%\begin{eqnarray}
%\mathbf{A} &=&
%\begin{bmatrix}
%   \mathbf{y}_{p}     & \dots    & \mathbf{y}_{N-1}\\
%   \mathbf{y}_{p-1}   & \dots    & \mathbf{y}_{N-2}\\
%   \vdots             & \ddots   & \vdots\\
%   \mathbf{y}_{1}     & \dots    & \mathbf{y}_{N-p}\\
%   1                  & \dots    & 1 
%   \end{bmatrix}^\top \, ,\\
%\mathbf{B} &=&
%\begin{bmatrix}
%\quad \\
%\mathbf{y}_{p+1} &
%\dots & 
%\mathbf{y}_N \\
%\quad
%\end{bmatrix}^\top  \, ,\\
%\mathbf{X}&=&
%\left[
%\begin{array}{ccccc}
%    \quad  & \quad & \quad & \quad \\
%    \phi_1   & \cdots & \phi_p & \mathbf{c} \\  
%    \quad & \quad & \quad & \quad
%    \end{array}
%\right]^\top \, ,\\
%\mathbf{E} &=& 
%\begin{bmatrix}
%    \quad \\
%    \mathbf{\epsilon}_{p+1}  & 
%    \dots                & 
%    \mathbf{\epsilon}_N \\
%    \quad
%\end{bmatrix}^\top \, . 
%\end{eqnarray}

\begin{equation}
 \label{eq:vareq}
               \underbrace{ \begin{bmatrix}
               \quad \\
               \mathbf{y}_{p+1} &
               \mathbf{y}_{p+2} &
               \dots & 
               \mathbf{y}_N \\
               \quad
               \end{bmatrix}}_{\substack{ \mathbf{B}^\top\\l \times (N-p)}}   
= 
                \underbrace{\left[ 
                \begin{array}{ccccc}
                \quad & \quad & \quad & \quad & \quad \\
                \phi_1  & \phi_2 & \cdots & \phi_p & \mathbf{c} \\  
                \quad &\quad & \quad & \quad & \quad
               \end{array} 
               \right]}_{\substack{ \mathbf{X}^\top\\ l \times (l \times p + 1 )}}
\underbrace{\begin{bmatrix}
   \mathbf{y}_{p}  & \mathbf{y}_{p+1} & \dots    & \mathbf{y}_{N-1}\\
   \mathbf{y}_{p-1}  & \mathbf{y}_{p} & \dots    & \mathbf{y}_{N-2}\\
   \vdots        & \vdots   & \ddots   & \vdots\\
   \mathbf{y}_{1} & \mathbf{y}_{2}   & \dots    & \mathbf{y}_{N-p}\\
   1 & 1   & \dots    & 1 
   \end{bmatrix}}_{\substack{ \mathbf{A}^\top\\ (l\times p +1 )\times (N-p)}}
+
\underbrace{\begin{bmatrix}
                \quad \\
              \mathbf{\epsilon}_{p+1}  & 
              \mathbf{\epsilon}_{p+2}  & 
              \dots                & 
              \mathbf{\epsilon}_N \\
              \quad
             \end{bmatrix}}_{\substack{\mathbf{E}^\top\\l \times (N-p) }} 
\end{equation}
Equation~(\ref{eq:vareq})  can be solved using ordinary least squares estimation.

\subsection{Vector Error Correction}

VECM is a special form of a VAR model for I(1) variables that are also
cointegrated~\cite{banerjee1993}. It is obtained by replacing the form $\Delta
\mathbf{y}_t = \mathbf{y}_t - \mathbf{y}_{t-1}$ in equation (\ref{eq:var}).
VECM is expressed in terms of differences, has an error correction term and 
the following form:

\begin{equation}
 \label{eq:vec}
 \Delta \mathbf{y}_t = 
 \underbrace{ \Omega\mathbf{y}_{t-1}}_\text{Error correction term} + 
 \sum_{i=1}^{p-1}
\phi_i^* \Delta \mathbf{y}_{t-i}  + \mathbf{c} + \mathbf{\epsilon}_t \quad ,
\end{equation}

\noindent where coefficients matrices $\Omega$ and $\phi_i^*$ are
function of matrices $\phi_i$ (shown in equation (\ref{eq:var})) as follows:

\begin{eqnarray*}
\phi_i^* &: =& -\sum_{j=i+1}^{p} \phi_j \\
\Omega &: =& -(\mathbb{I}-\phi_1-\dots-\phi_p) 
\end{eqnarray*}

The matrix $\Omega$ has the following properties~\cite{johansen1995}:
\begin{itemize}
\item If $\Omega = 0$ there is no cointegration
\item If $rank(\Omega)=l$ i.e full rank, then the time series are not
I(1) but stationary
\item If $rank(\Omega)=r,\quad 0 < r < l$ then, there is cointegration
and the matrix $\Omega$ can be expressed as $\Omega =
\alpha \beta^\top$, where $\alpha$ and $\beta$ are $(l \times r)$
matrices and $rank(\alpha)=rank(\beta)=r$.

The columns of $\beta$ contains the cointegration vectors and the rows of
$\alpha$ correspond with the adjusted vectors. $\beta$ is obtained by Johansen
procedure~\cite{johansen1988} whereas $\alpha$ has to be determined as a
variable in the VECM.

It is worth noticing that the factorization of the matrix $\Omega$ is not
unique since for any $r \times r$ nonsingular matrix $H$ we have:

\begin{eqnarray*}
\alpha \beta^\top &=& \alpha \mathbf{HH^{-1}} \beta^\top\\
&=&(\alpha\mathbf{H})(\beta(\mathbf{H}^{-1})^\top)^\top \\
&=& \alpha^*(\beta^*)^\top
\end{eqnarray*}

\noindent whith $\alpha^* = \alpha\mathbf{H}$ and $\beta^* =
\beta(\mathbf{H}^{-1})^\top$.

\end{itemize}

If cointegration exists, then equation (\ref{eq:vec}) can be written as follows:

\begin{equation}
 \label{eq:vecfull}
 \Delta \mathbf{y}_t = \alpha \beta^\top\mathbf{y}_{t-1} 
 + \sum_{i=1}^{p-1} \phi_i^*\Delta
\mathbf{y}_{t-i}  + \mathbf{c} + \mathbf{\epsilon}_t \quad ,
\end{equation}

\noindent which is a VAR model but for time series differences.

VECM matricial form is the following:

%\begin{eqnarray}\label{eq:vecmatrix}
% \mathbf{A}&=&
%   \begin{bmatrix} 
%   \beta^\top \mathbf{y}_{p} & 
%   \cdots & \beta^\top \mathbf{y}_{N-1} \\
%   \Delta \mathbf{y}_p  & \cdots 
%   &\Delta\mathbf{y}_{N-1} \\ 
%   \vdots  & \ddots & \vdots \\
%   \Delta\mathbf{y}_2  & \cdots 
%   & \Delta \mathbf{y}_{N-p+1} \\
%   1 & \cdots & 1 
%   \end{bmatrix}^\top \label{eq:vecA} \, ,\\
%\mathbf{B} & = &
% \begin{bmatrix}
% \quad\\
%  \Delta \mathbf{y}_{p+1} & 
%  \dots &
%  \Delta \mathbf{y}_N \label{eq:vecB}\\
%  \quad
% \end{bmatrix}^\top \, ,\\
%\mathbf{X}&=&
%  \begin{bmatrix}
%   \quad \\
%   \alpha & \phi_1^* & \cdots & \phi_{p-1}^* & \mathbf{c} \\  
%   \quad
%   \end{bmatrix}^\top \, ,\\
%\mathbf{E} &=&
%\begin{bmatrix}
%   \quad \\
%   \mathbf{\epsilon}_{p+1} &
%   \dots &
%   \quad &\mathbf{\epsilon}_N \, \\ \quad
%\end{bmatrix}^\top 
%\end{eqnarray}
%

\begin{equation} \label{eq:vecmatrix}
\underbrace{
               \begin{bmatrix}
               \quad\\
                \mathbf{\Delta y}_{p+1} & 
                \dots &
                \mathbf{\Delta y}_N \\
                \quad
               \end{bmatrix}
               }_{\substack{\mathbf{B}^\top\\ l \times (N-p) }} =
   \underbrace{
    \begin{bmatrix}
     \quad \\
     \alpha & \phi_1^* & \cdots & \phi_{p-1}^* & \mathbf{c} \\  
     \quad
     \end{bmatrix} 
     }_{\substack{ \mathbf{X}^\top\\ l \times (l\times p +1)}}
\underbrace{\begin{bmatrix} 
   \beta^\top \mathbf{y}_{p} & 
   \cdots & \beta^\top \mathbf{y}_{N-1} \\
   \mathbf{\Delta y}_p  & \cdots 
   & \mathbf{\Delta y}_{N-1} \\ 
   \vdots  & \ddots & \vdots \\
   \mathbf{\Delta y}_2  & \cdots 
   & \mathbf{\Delta y}_{N-p+1} \\
   1 & \cdots & 1 
   \end{bmatrix}}_{\substack{\mathbf{A}^\top \\ (l \times p +1) \times (N-p) }}
+
\underbrace{\begin{bmatrix}
              \quad \\
              \mathbf{\epsilon}_{p+1} &
              \dots &
              \quad &\mathbf{\epsilon}_N \\ \quad
             \end{bmatrix}}_{\substack{\mathbf{E}^\top\\ l \times (N-p) }} 
\end{equation}

VAR and VECM parameters shown in equations (\ref{eq:vareq},\ref{eq:vecmatrix}) can be solved
using standard regression techniques, such as ordinary least squares (OLS).

\subsection{Unit Root Tests}
In order to use VECM the time series are required to be I(1) which is detected
using a unit root test. The Dickey-Fuller (ADF) test ~\cite{dickey1979} is a
unit root test where the null hypothesis is that the
variable has a unit root.

If we have a variable with the form:
\[
\mathbf{x}_t = \alpha \mathbf{x}_{t-1} + \epsilon_t
\]

\noindent and if we substract $\mathbf{x}_{t-1}$ from both sides we get:

\begin{equation}
\label{eq:DF}
\Delta \mathbf{x}_t = (\alpha-1) \mathbf{x}_{t-1} + \epsilon_t
\end{equation}

Dickey fuller test consist on determine if $\alpha$ equals 1. The alternative
hypothesis is $\alpha < 1$ which means that $\mathbf{x}_t$ is a stationary
process.

In order to include more than 1 lagged variable the Augmented Dickey-Fuller
test was developed and their tabulated critical values were obtained
by~\cite{mackinnon2010}. The number of lags is often selected using Schwartz
criterion.

\subsection {Cointegration Tests}

Johansen method~\cite{johansen1988} allows to find cointegration among several
I(1) time series. This test is more general than the
Engle-Granger~\cite{engle1987} test which only allow to find cointegration
between two time series.  Johansen procedure takes the form of a
likelihood-ratio test which determines $r$ cointegration relationships. The
method provides two tests based on an eigenvector decomposition of the matrix
$\Omega = \alpha \beta^\top$ on equation~\ref{eq:vec} called trace and maximal
eigenvalue statistic.  Once the number of cointegration number is determined,
$r$ more significative eigenvectors provided by Johansen procedure are used to
get the cointegration vector matrix $\beta^\top$.  Critical values for both
tests are limited to 12 and 5 variables in the trace and maximal eigenvalue
statistic respectively.

The trace test null hypothesis is $H_0: r = r_0$ against $H_1: r > r_0$ and the
maximum eigenvalue test null hypothesis is $H_0: r = r_0$ against $H_1: r =
r_0+1$ starting with $r_0 = 0$.  In the latest test, if we start with $r_0=0$
and $H_0$ is rejected $H_1: r = 1$ will be accepted which is not true if there
are more than one cointegration relationship, this is why this test is less
powerful and the trace test is commonly used.

\subsection{Ordinary Least Squares method}

When $\mathbf{A}$ is singular, solution to equation~(\ref{eq:vecmatrix}) is given
by the ordinary least squares (OLS) method. OLS consists of minimizing the sum
of squared errors or equivalently minimizing the following expression:

\begin{equation}
\label{eq:regressionproblem}
\underset{\mathbf{X}}{\text{min}} \quad \| \mathbf{A}\mathbf{\mathbf{X}} - \mathbf{B} \|_2^2
\end{equation}

\noindent for which the solution $\hat{\mathbf{X}}$ is well-known:

\begin{equation}
\label{eq:MP}
\hat{\mathbf{X}}=\mathbf{A}^{\!\!+}\,\mathbf{B}
\end{equation}

\noindent where $\mathbf{A}^{\!\!+}$ is the Moore-Penrose pseudo-inverse
which can be written as follows:

\begin{equation}
\label{eq:pseudoinverse}
\mathbf{A}^{\!\!+}= (\mathbf{A}^{\!\!\top} \mathbf{A})^{-1}\mathbf{A}^{\!\!\top} \, .
\end{equation}

\begin{equation}
    \label{eq:compactsvd}
    \underset{m \times n}{\mathbf{A}}=
    \underset{m \times k}{\mathbf{U_1}} \enskip
    \underset{k \times k}{\Sigma_1} \enskip
    \underset{k \times n}{\mathbf{V}_1^{\top}} \, ,
\end{equation}

\begin{equation}
\label{eq:pseudoinversesvd}
\mathbf{A}^{\!\!+} = \mathbf{V}_1 \Sigma_1^{-1} \mathbf{U}_1^\top \, .
\end{equation}

%Descomentar en la tesis
\textbf{Demo}\quad

Since $\mathbf{A}$ is singular the problem shown in
equation~(\ref{eq:regressionproblem}) has not solution, the minimum norm given
by equation~(\ref{eq:MP}) is obtained by solving the equivalent problem:

\begin{equation*}
\label{eq:proyectorsol}
\mathbf{A \hat{X} = PB} 
\end{equation*}

Since $\mathbf{V} = [\underset{(n \times k)}{\mathbf{V_1}} |
\underset{(n \times k)}{\mathbf{V_2}}]$ and $\mathbf{V_1^\top V_2 =
0}$ we can express $\mathbf{\hat{X}} = \mathbf{V_1 x_1 + V_2 x_2}$
with $\mathbf{x_2=0}$ because $\mathbf{\hat{X}}$ lives in the
$\text{Row}(\mathbf{A})$ given by $\mathbf{V_1}$, so we have:

\begin{eqnarray*}
\mathbf{A \hat{X}} &=& \mathbf{PB} \\
\mathbf{U_1 \Sigma_1 V_1^\top \hat{X}} &=& \mathbf{U_1 U_1^\top B} \\
\mathbf{ V_1^\top \hat{X}} &=&  \mathbf{\Sigma_1^{-1} U_1^\top B} \\ 
\mathbf{ V_1^\top V_1 x_1} &=& \mathbf{\Sigma_1^{-1}
U_1^\top B} \\
\mathbf{x_1}&=& \mathbf{\Sigma_1^{-1} U_1^\top B}
\end{eqnarray*}

\noindent from this result we can obtain $\mathbf{\hat{X}}$ and
therefore the pseudo-inverse expression:

\begin{eqnarray*}
\mathbf{\hat{X}} &=& \mathbf{V_1 x_1} \\
                &=& \mathbf{V_1 \Sigma_1^{-1} U_1^\top B} \\
\mathbf{A^+} &=& \mathbf{V_1 \Sigma_1^{-1} U_1^\top} \, .
\end{eqnarray*}

\subsection{Ridge Regression}\label{sec:RR}

Ridge regression (RR) introduces a ridge estimation to avoid the
singularity of the matrix $\mathbf{A}^\top \mathbf{A}$
equation~(\ref{eq:pseudoinverse}). RR solves the following
optimization problem:

\begin{equation}
\label{eq:RRproblem}
\underset{\mathbf{X}}{\text{min}} \quad \|
\mathbf{A}\mathbf{\mathbf{X}} - \mathbf{B} \|_2^2 +\lambda \|
\mathbf{\mathbf{X}}\|_2^2 
\end{equation}

\noindent where $\lambda$ is a regularization parameter.

The RR optimal solution $\mathbf{X}(\lambda)$ is also well-known:

\begin{equation}
\label{eq:optsolRR}
\mathbf{X}(\lambda)=(\mathbf{A}^\top \mathbf{A}+ \lambda
\mathbb{I})^{-1}\mathbf{A}^\top \mathbf{B} \, . 
\end{equation}

@@

To solve equation~(\ref{eq:optsolRR}) is equivalent to solve the following
optimization problem:

@@

\begin{equation*}
	\mathbf{C}=\begin{bmatrix} \quad &\mathbf{A}& \quad \\ \hdashline \quad
& \lambda\mathbb{I}&\quad  \end{bmatrix} \quad \text{and} \quad
	\mathbf{F}=\begin{bmatrix} \quad &\mathbf{B}& \quad \\ \hdashline \quad
& \mathbf{0}&\quad  \end{bmatrix} \, .
\end{equation*}

Applying equation~(\ref{eq:MP}) and considering that $\mathbf{C}^\top
\mathbf{C} = \mathbf{A}^\top \mathbf{A} + \lambda \mathbb{I}$ and 
$\mathbf{C}^\top \mathbf{F}=\mathbf{A}^\top \mathbf{B} $ we have:

\begin{eqnarray*}
\mathbf{X}(\lambda)&=&(\mathbf{C}^\top
\mathbf{C})^{-1}\mathbf{C}^\top \mathbf{F} \\
&=& (\mathbf{A}^\top \mathbf{A} + \lambda \mathbb{I})^{-1} \mathbf{A}^\top \mathbf{B}
\end{eqnarray*}

\subsection{Online Ridge regression}

If we represent matrices $\mathbf{A}$ and $\mathbf{B}$ such as:
